---
title: "Homework 1"
author: "FIRSTNAME LASTNAME"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Instructions

**Exercises:** 1-5 (Pgs. 6-7); 1-2, 5 (Pg. 12); 1-5 (Pgs. 20-21); Open Response

**Submission:** Submit via an electronic document on Sakai. Must be submitted as a HTML file generated in RStudio. All assigned problems are chosen according to the textbook *R for Data Science*. You do not need R code to answer every question. If you answer without using R code, delete the code chunk. If the question requires R code, make sure you display R code. If the question requires a figure, make sure you display a figure. A lot of the questions can be answered in written response, but require R code and/or figures for understanding and explaining.

```{r, include=FALSE}
library(tidyverse)
library(ggplot2)
```

# Chapter 1 (Pgs. 6-7)

##  Exercise 1
```{r}
ggplot(data=mpg)
```

I see absolutely nothing. There is just a blank space for a graph. Why am I even doing this nonsense?

## Exercise 2
```{r}
dim(mpg)
nrow(mpg)
ncol(mpg)
```
There are 234 rows and 11 columns in the dataset **mpg**.

##  Exercise 3
```{r, eval=FALSE}
?mpg
```

```{r}
unique(mpg$drv)
```


The variable **drv** is a factor variable that takes the following values:

- "f" = front-wheel drive
- "r" = rear-wheel drive
- "4" = 4-wheel drive

##  Excercise 4
```{r}
ggplot(data=mpg,aes(x=hwy,y=cyl)) +
  geom_point() + 
  xlab("Highway Miles Per Gallon") +
  ylab("Number of Cylinders")
```


##  Excercise 5
```{r}
ggplot(data=mpg,aes(x=class,y=drv)) + 
  geom_point() + 
  xlab("Type of Car") +
  ylab("Type of Drive")
```

Scatter plots are not meant to visualize the relationship between two categorical/qualitative variables. 

# Chapter 1 (Pg. 12)

##  Exercise 1
```{r}
#incorrect
ggplot(data = mpg) +
 geom_point(
 mapping = aes(x = displ, y = hwy, color = "blue")
 )
#correct
ggplot(data = mpg) +
 geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
```

What’s gone wrong with this code? Why are the points not blue?
In order to change the color of the dots we need to set an aesthetic manually. Setting manually requires assigning a color outside of the aes function. 

##  Exercise 2

Which variables in mpg are categorical? Which variables are continuous? (Hint: type ?mpg to read the documentation for the dataset.) How can you see this information when you run mpg?

Categorical Variables: manufacturer, model, year, trans, drv, fl, class, cyl
Continuous Variables: hwy, cty, displ
When you run the mpg data set you are able to see the individual data points. Data points that are categorical are fixed where as continous variables arnt. 


```{r}
?mpg
```

##  Exercise 5

What does the stroke aesthetic do? What shapes does it work with? (Hint: use ?geom_point.)
The stroke aesthetic is used to modify the width of borders for shapes that include borders. 


```{r}
?geom_point
```

# Chapter 1 (Pgs. 20-21)

##  Exercise 1
Geom Definition from notes: A geom is the geometrical object that a plot uses to represent data.

1. What geom would you use to draw a line chart? A boxplot? Ahistogram? An area chart?
Line chart: Use Line Geoms
Boxplot: Use boxplot Geoms
Histogram: Use histogram Geom
Area Chart: Use area geom.
Scatterplot: Point Geoms



##  Exercise 2

```{r}
# . Run this code in your head and predict what the output will look like. Then, run the code in R and check your predictions:

ggplot(
 data = mpg,
 mapping = aes(x = displ, y = hwy, color = drv)
) +
 geom_point() +
 geom_smooth(se = FALSE)

```

Prediction: This will produce a plot of the mpg data that includes both the point and smooth geoms. The plot will have disply as its x varible and hwy as its y variable. I think that the point geom will have different colored pointed based on the different classes of drv recorded. I think the smooth geom will be one solid color line due to the code se = FLASE (iffy about this part). 

# Check prediction: I was correct about everything except the output of the smooth geom. Instead of being one solid colored line it was three lines with different colors that correspond to different classes of drv. I am still unse what se = FALSE does.I tried testing what it does by removing se  = FALSE from the code but I did not see a change in the plot produced when run. 

##  Exercise 3

 What does show.legend = FALSE do? What happens if you remove it? Why do you think I used it earlier in the chapter?
# show.legend = FLASE removes the legend produced on the plot when assigning a variable to different colors inside aesthetic. If you remove this I would assume the legend is included on the graph. Maybe the author removed it earlier in the chapter the legend made the plot too full or busy. 


##  Exercise 4

What does the se argument to geom_smooth() do?
According to the help center the se argument to geom smooth "display confidence interval around smooth? (TRUE by default, see level to control.)". se stands for standard error. 


```{r}
?geom_smooth
```

##  Exercise 5

I don't know if they will look different. Let me check.

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()

ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))
```

They do not look different. I am incredibly surprised.

# Open Response

For this exercise, use the **diamonds** dataset in the tidyverse. Use `?diamonds` to get more information about the dataset.

## Step 1: Select 1 numeric variable and 2 categorical variables. Create a graphic using `geom_boxplot()` and `facet_wrap` to illustrate the empirical distributions of the sample.
```{r}
?facet_wrap
head(diamonds)
```

```{r}

ggplot(data = diamonds) + geom_boxplot(mapping = aes(x = price)) +
  facet_grid(cut ~ clarity)


```

## Step 2: Choose 2 numeric variables and 2 categorical variables and creatively illustrate the relationship between all the variables.

```{r}
ggplot(data=diamonds, mapping = aes(x = carat, y = price)) + 
  geom_point(mapping = aes(shape=cut,color=color))

```

